### 2020-06-10 FDT2020/FDT.tex
somefixes
Initial abstract and some comments

@@

\usepackage{mathrsfs}

\usepackage{epigraph}
\usepackage{thm-restate}
\usepackage{cleveref}

@@

\begin{abstract}

@@

\input{Introduction}
\input{main section}

>>>>

\newcommand{\cindy}[1]{\textcolor{green!60!black}{(Cindy: #1)}}
\newcommand{\arash}[1]{\textcolor{blue}{(Arash: #1)}}
\newcommand{\bob}[1]{\textcolor{red}{(Bob: #1)}}
% To drop all comments, just comment out \iffalse and \fi
\iffalse
\renewcommand{\cindy}[1]{}
\renewcommand{\arash}[1]{}
\renewcommand{\bob}[1]{}
\fi

@@

\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage{epigraph}
\usepackage{cleveref}

@@

\begin{abstract}
We present a new algorithm, Fractional Decomposition Tree (FDT) for finding a feasible solution for an integer program (IP) where all variables are binary. FDT runs in polynomial time and is guaranteed to find a feasible integer solution provided the integrality gap is bounded. The algorithm gives a construction for Carr and Vempala's theorem that any feasible solution to the IP's linear-programming relaxation, when scaled by the instance integrality gap, dominates a convex combination of feasible solutions. FDT is also a tool for studying the integrality gap of IP formulations.  We demonstrate that with experiments studying the integrality gap of two problems: optimally augmenting a tree to a 2-edge-connected graph and finding a minimum-cost 2-edge-connected multi-subgraph (2EC). We also give a simplified algorithm, Dom2IP, that more quickly determines if an instance has an unbounded integrality gap. We show that FDT's speed and approximation quality compare well to that of feasibility pump on moderate-sized instances of the vertex cover problem. For a particular set of hard-to-decompose fractional 2EC solutions, FDT always gave a better integer solution than the best previous approximation algorithm (Christofides).

@@

\input{Introduction}
\input{main-section}

### 2020-06-10 FDT2020/Introduction.tex
Initial abstract and some comments

In combinatorial optimization the aim is to find the optimal solution in a discrete and
usually finite yet large set of solutions. For many specific combinatorial optimization problems such a solution can be found efficiently. For many others, finding optimal or in many cases near optimal solutions is NP-hard. A common approach to deal with such problems is relaxing the discrete solution set into a continuous set, where the optimization problem becomes tractable. Obtaining feasible solutions by means of such a relaxation requires an additional step of rounding the potentially fractional solution of the continuous relaxation into integer solutions.

In this paper, our focus is on linear relaxation of combinatorial
optimization problems. Combinatorial optimization was pioneered by Edmonds even before efficient algorithms for solving linear programming problems where introduced by Khachiyan \cite{Khachiyan} and later by Karmarkar \cite{Karmarkar1984}. For problems such as the \textsc{Minimum Cost Spanning Tree Problem} there are linear programming relaxations whose basic feasible solutions coincide with  integral solutions, i.e. spanning trees. For other problems the value of the linear programming relaxation provides a bound (lower bound for a minimization problem and upper bound for a maximization problem) on the optimal solution. A common and successful approach is to round these (potentially) fractional solutions into integer solutions for the combinatorial optimization problem at hand. The Integrality gap of a linear relaxation of an integer programming problem is the worst case ratio between the objective values of the discrete problem and the continuous problem. Equivalently, the integrality gap of the linear programming relaxation is a limit to the rounding approach: rounding a fractional solution into an integer solution incurs
a multiplicative cost proportional to the integrality gap. In this dissertation we study integrality gaps for different combinatorial optimization problems and introduce new rounding algorithms that imply bounds on their respective integrality gaps.

@@

Let $S$ denote the set of feasible solutions to a combinatorial optimization problem. For instance, for many problems in network optimization, set $S$ is a subset of $ \{0,1\}^n$ where each coordinate of a point in $S$ indicates the absence or presence of the corresponding edge in a solution, and $n$ is the number of edges in the network. Suppose set $S$ can be described as $S=\{x\in \mathbb{Z}^n: Ax\geq b, x\geq 0 \}$ for some $A\in \mathbb{R}^{m\times n}$ and $b\in \mathbb{R}^m$. (Pure) Integer Programming (IP) asks for $\min_{x\in S}cx$ for some $c\in \mathbb{R}^n$. Integer programming is NP-hard and in fact, it is even NP-complete to decide whether set $S$ is empty or not \cite{GareyJohnson}. The convex hull of $S$ denoted by $\conv(S)$ is the minimal convex set containing $S$ and can be formulated as follows.
\begin{equation*}
\conv(S) =\{\sum_{i=1}^{k} \lambda_i x^i: x^i \in S \text{ for } i =1,\ldots,k,\; \lambda_i \geq 0 \text{ for } i = 1,\ldots,k \text{, and } \sum_{i=1}^{k}\lambda_i = 1\}.
\end{equation*}

@@

Let $I=(A,b)$ denote an instance. Then $S(I)$ and $P(I)$ denote $S(A,b)$ and $P(A,b)$, respectively. Given a linear objective function $c$, recall that an IP is $\min \;\{cx:\; x \in S(I)\}$. It is  NP-hard even to determine if an IP instance has a feasible solution~\cite{GareyJohnson}. However, intelligent branch-and-bound strategies allow commercial and open-source MILP solvers to give exact solutions (or nearoptimal solutions with a provable bound) to many specific instances of NP-hard combinatorial optimization problems.

Relaxing the integrality constraints gives the polynomial time solvable linear programming relaxation: $\min \;\{cx:\;x\in P(I) \}$.  The optimal value of this linear program (LP), denoted $z_{\lp}(I,c)$, is a lower bound on the optimal value for the IP, denoted $z_{\IP}(I,c)$. The solutions can also provide some useful global structure, even though the fractional values might not be directly meaningful.

Many researchers (see \cite{davids,vazirani}) have developed polynomial time LP-based approximation algorithms that find solutions for special classes of IPs whose cost are provably smaller than $C\cdot z_{LP}(I,c)$. The approximation factor $C$ can be a constant or depend on the input parameters of the IP, e.g. $O(\log(n))$ where $n$ is the number of variables in the formulation of the IP (the dimension of the problem). However, for many combinatorial optimization problems there is a limit to such techniques based on LP relaxations, represented  by the {integrality gap} of the IP formulation. Recall that integrality gap $g(I)$ for instance $I$ is defined to be $g(I)= \max_{c\geq 0}\frac{z_{IP}(I,c)}{z_{LP}(I,c)}$. An example of an instance specific integrality gap is the integrality gap of the subtour elimination relaxation for the 2-edge-connected spanning multigraph problem on $n$ vertices. The instance is the complete graph on $n$ vertices. Alexander et al. \cite{alexander2006integrality} showed the instance specific integrality gap of the subtour elimination relaxation for the 2-edge-connected multigraph problem for instances of the problem with $n= 10$ is at most $\frac{7}{6}$.

@@

Our methods apply theory connecting integrality gaps to sets of feasible solutions. Instances $I$ with  $g(I)=1$ has $P(I)=\conv(S(I))$, the convex hull of the lattice of feasible points. In this case, $P(I)$ is an \textit{integral} polyhedron. The spanning tree polytope of graph $G$, $\st(G)$, and the perfect-matching polytope of graph $G$, $\permat(G)$, have this property (\cite{Edmonds2003,edmondsPM}). For such problems there is an algorithm to express vector $x\in P(I)$ as a convex combination of points in $S(I)$ in polynomial time \cite{cons-cara}.
\begin{proposition}\label{cara}
	If $g(I)=1$, then for $x\in P(I)$ there exists $\theta \in [0,1]^k$, where $\sum_{i=1}^{k}\theta_i =1$ and $\tilde{x}^i\in S(I)$ for $i\in [k]$ such that $\sum_{i=1}^{k}\theta_i \tilde{x}^i\leq x$. Moreover, we can find such a convex combination in polynomial time.
\end{proposition}

@@

We give a general approximation framework for solving binary IPs.  Consider the set of point described by sets $S(I)$ and $P(I)$ as in (\ref{S}) and (\ref{P}), respectively. Assume in addition that $S(I),P(I)\subseteq [0,1]^n$. For a vector $x\in \mathbb{R}_{\geq 0}^n$ such that $x\in P(I)$, let $\spp(x)= \{i \in [n]: x_i \neq 0\}$. For an integer $\beta$ let $\{\beta\}^n$ be the vector $y\in \mathbb{R}^n$ with $y_i=\beta$ for $i\in [n]$.

@@

\subsection{Experiments.} Although the bound guaranteed in both Theorems \ref{binaryFDT} and \ref{FDT2EC} are very large, we show that in practice, the algorithm works very well for network design problems described above. We show how one might use FDT to investigate the integrality gap for such well-studied problems.

@@

In the \textsc{minimum vetex cover problem (VC)} we are given a graph $G=(V,E)$ and $c\in \mathbb{R}^E_{\geq 0}$. A subset of $U$ of $V$ is a \textit{vertex cover} if for $e\in E$ at least one endpoint of $e$ is in $U$. The goal in VC is to find the minimum cost vertex cover. The linear programming relaxation for VC is
\begin{equation}
\min \{cx \; : \; x_u + x_v \geq	 1 \text{ for } e=uv \in E, \; x\in [0,1]^{V}\}.
\end{equation}
The integrality gap of this formulation is exactly 2. It is shown that it is UG-hard to approximte VC within any factor sctrictly better than 2 \cite{UGhardVC}. We compare  FDT and the feasbility pump heuristic \cite{fp1} on the PACE 2019\footnote{ Parameterized Algorithms and Computational Experiments Challenge 2019: \url{https://pacechallenge.org/2019/}} challenge test cases \cite{PACE} and compare its performance with 
\subsubsection{Tree augmentation problem}
In the \textsc{Tree Augmentation Problem (TAP)} we are given a  graph $G=(V,E)$, a a spanning tree $T$ of $G$. We also have a cost vector $c\in \mathbb{R}^{E\setminus T}_{\geq 0}$. A subset $F$ of $E\setminus T$ is called a \textit{feasible augmentation} if $(V,T\cup F)$ is a 2-edge-connected graph. In TAP we seek the minimum cost feasible augmentation. The natural linear programming relaxation for TAP is 
\begin{equation}
\min \{cx\; : \; \sum_{\ell \in \cov(e)} x_{\ell} \geq 1 \text{ for } e\in T, \; x\in [0,1]^{E\setminus T}\}.
\end{equation}
where $\cov(e)$ is set of edges $\ell \in E\setminus T$ such that $e$ is in the unique cycle of $T\cup \{\ell\}$. We call the LP above the cut-LP. The integrality gap of the cut-LP is known to be between $\frac{3}{2}$ \cite{32gaptap} and $2$ \cite{FJ81}. We create random fractional extreme points of the cut-LP and round them using FDT. For the instances that we create the blow-up factor is always below $\frac{3}{2}$ providing an upper bound for such instances.

@@

Consider a graph $G=(V,E)$. A \textit{Carr-Vempala point} $x\in \mathbb{R}^E$ is a fractional point in $\subtour(G)$ where the edges $e$ with $0<x_e<1$ form a Hamiltonian cycle of $G$. Carr and Vempala \cite{Carr2004} showed that $g(\2ec)$ for instances where the optimal solution to $\min_{x\in \subtour(G)}cx$ is a Carr-Vempala point. We show that the integrality gap is at most $\frac{6}{5}$ for Carr-Vempala points with at most 12 vertices on the Hamiltonian cycle formed by the fractional edges. For Carr-Vempala points we assume that 1-edges are replaced by long paths of 1-edges making these points into potentially harder to round instances.

>>>>

\cindy{An original comment: There's material in this section (e.g. general discussion of combintatorial optimization) that's likely unnecessary for the journal paper. We probably want to start with what is now Section 2, but have some opening introduction giving the big picture.  Remember when we first mention rounding to be clear this is not numerical rounding. TODO: Add English statement for things currently only given in math. Add definitions of problems so reviewers and readers don't have to look them up.}
In combinatorial optimization the aim is to find the optimal solution in a discrete and
usually finite yet large set of solutions. For many specific combinatorial optimization problems such a solution can be found efficiently. For many others, finding optimal or in many cases near optimal solutions is NP-hard. A common approach to deal with such problems is relaxing the discrete solution set into a continuous set, where the optimization problem becomes tractable. Obtaining feasible solutions by means of such a relaxation requires an additional step of rounding the potentially fractional solution of the continuous relaxation into integer solutions.

\cindy{TODO: remove ``recall that'' for anything we haven't already defined in this paper.  Even though it's likely anyone reading this will care about IPs and integrality gaps, we should not assume they already know things we define the way we define them.}
In this paper, our focus is on linear relaxation of combinatorial
optimization problems. Combinatorial optimization was pioneered by Edmonds even before efficient algorithms for solving linear programming problems where introduced by Khachiyan \cite{Khachiyan} and later by Karmarkar \cite{Karmarkar1984}. For problems such as the \textsc{Minimum Cost Spanning Tree Problem} there are linear programming relaxations whose basic feasible solutions coincide with  integral solutions, i.e. spanning trees. For other problems the value of the linear programming relaxation provides a bound (lower bound for a minimization problem and upper bound for a maximization problem) on the optimal solution. A common and successful approach is to round these (potentially) fractional solutions into integer solutions for the combinatorial optimization problem at hand. The Integrality gap of a linear relaxation of an integer programming problem is the worst case ratio between the objective values of the discrete problem and the continuous problem. Equivalently, the integrality gap of the linear programming relaxation is a limit to the rounding approach: rounding a fractional solution into an integer solution incurs
a multiplicative cost proportional to the integrality gap. In this dissertation we study integrality gaps for different combinatorial optimization problems and introduce new rounding algorithms that imply bounds on their respective integrality gaps.

@@

Let $S$ denote the set of feasible solutions to a combinatorial optimization problem. For instance, for many problems in network optimization, set $S$ is a subset of $ \{0,1\}^n$ where each coordinate of a point in $S$ indicates the absence or presence of the corresponding edge in a solution, and $n$ is the number of edges in the network. Suppose set $S$ can be described as $S=\{x\in \mathbb{Z}^n: Ax\geq b, x\geq 0 \}$ for some $A\in \mathbb{R}^{m\times n}$ and $b\in \mathbb{R}^m$. (Pure) Integer Programming (IP) asks for $\min_{x\in S}cx$ for some $c\in \mathbb{R}^n$. Integer programming is NP-hard and in fact, it is even NP-complete to decide whether set $S$ is empty or not \cite{GareyJohnson}. The convex hull of $S$ denoted by $\conv(S)$ is the minimal convex set containing $S$ and can be formulated as follows.
\begin{equation*}
\conv(S) =\{\sum_{i=1}^{k} \lambda_i x^i: x^i \in S \text{ for } i =1,\ldots,k,\; \lambda_i \geq 0 \text{ for } i = 1,\ldots,k \text{, and } \sum_{i=1}^{k}\lambda_i = 1\}.
\end{equation*}
\cindy{Since most canonical formulations for IP/LP express variables bounds separately, we should mention here that they are rolled into matrix $A$.}

@@

Let $I=(A,b)$ denote an instance. Then $S(I)$ and $P(I)$ denote $S(A,b)$ and $P(A,b)$, respectively. Given a linear objective function $c$, recall that an IP is $\min \;\{cx:\; x \in S(I)\}$. It is  NP-hard even to determine if an IP instance has a feasible solution~\cite{GareyJohnson}. However, intelligent branch-and-bound strategies allow commercial and open-source MILP solvers to give exact solutions (or near-optimal solutions with a provable bound) to many specific instances of NP-hard combinatorial optimization problems.

Relaxing the integrality constraints gives the polynomial-time-solvable linear programming relaxation: $\min \;\{cx:\;x\in P(I) \}$.  The optimal value of this linear program (LP), denoted $z_{\lp}(I,c)$, is a lower bound on the optimal value for the IP, denoted $z_{\IP}(I,c)$. The solutions can also provide some useful global structure, even though the fractional values might not be directly meaningful.

Many researchers (see \cite{davids,vazirani}) have developed polynomial time LP-based approximation algorithms that find solutions for special classes of IPs whose cost are provably smaller than $C\cdot z_{LP}(I,c)$. The approximation factor $C$ can be a constant or depend on the input parameters of the IP, e.g. $O(\log(n))$ where $n$ is the number of variables in the formulation of the IP (the dimension of the problem). However, for many combinatorial optimization problems there is a limit to such techniques based on LP relaxations, represented  by the {integrality gap} of the IP formulation. Recall that integrality gap $g(I)$ for instance $I$ is defined to be $g(I)= \max_{c\geq 0}\frac{z_{IP}(I,c)}{z_{LP}(I,c)}$. \cindy{Does the integrality gap always require positive objective coefficients?  Or is that an extra restriction in our definition?  Is there a reference that has that restriction?} An example of an instance-specific integrality gap is the integrality gap of the subtour-elimination relaxation for the 2-edge-connected spanning multigraph problem on $n$ vertices. \cindy{We might need to define the problem or have a forward pointer.} The instance is the complete graph on $n$ vertices. Alexander et al. \cite{alexander2006integrality} showed the instance specific integrality gap of the subtour elimination relaxation for the 2-edge-connected multigraph problem for instances of the problem with $n= 10$ is at most $\frac{7}{6}$.

@@

Our methods apply theory connecting integrality gaps to sets of feasible solutions. Instances $I$ with  $g(I)=1$ has $P(I)=\conv(S(I))$, the convex hull of the lattice of feasible points. In this case, $P(I)$ is an \textit{integral} polyhedron. The spanning tree polytope of graph $G$, $\st(G)$, and the perfect-matching polytope of graph $G$, $\permat(G)$, have this property (\cite{Edmonds2003,edmondsPM}). For such problems there is an algorithm to express vector $x\in P(I)$ as a convex combination of points in $S(I)$ in polynomial time \cite{cons-cara}.
\cindy{Should this proposition just be a corollary of the Carr-Vempala theorem?  Are we going to describe how to do it?  If so, there should be a forward pointer to that discussion.}
\begin{proposition}\label{cara}
	If $g(I)=1$, then for $x\in P(I)$ there exists $\theta \in [0,1]^k$, where $\sum_{i=1}^{k}\theta_i =1$ and $\tilde{x}^i\in S(I)$ for $i\in [k]$ such that $\sum_{i=1}^{k}\theta_i \tilde{x}^i\leq x$. Moreover, we can find such a convex combination in polynomial time.
\end{proposition}

@@

We give a general approximation framework for solving binary IPs.  Consider the set of point described by sets $S(I)$ and $P(I)$ as in (\ref{S}) and (\ref{P}), respectively. Assume in addition that $S(I),P(I)\subseteq [0,1]^n$. \cindy{State somewhere that we can allow continuous variables as long as they are not in the objective function. (We will explain later how to extend to this case).} For a vector $x\in \mathbb{R}_{\geq 0}^n$ such that $x\in P(I)$, let $\spp(x)= \{i \in [n]: x_i \neq 0\}$. For an integer $\beta$ let $\{\beta\}^n$ be the vector $y\in \mathbb{R}^n$ with $y_i=\beta$ for $i\in [n]$.

@@

\subsection{Experiments.} Although the bound guaranteed in both Theorems \ref{binaryFDT} and \ref{FDT2EC} are very large, we show that in practice, the algorithm works very well for network design problems described above. We show how one might use FDT to investigate the integrality gap for such well-studied problems. \cindy{TODO: forward pointers to the sections with more details.}

@@

In the \textsc{minimum vetex cover problem (VC)} we are given a graph $G=(V,E)$ and $c\in \mathbb{R}^E_{\geq 0}$. A subset of $U$ of $V$ is a \textit{vertex cover} if for $e\in E$ at least one endpoint of $e$ is in $U$. The goal in VC is to find the minimum cost vertex cover. The linear programming relaxation for VC is
\begin{equation}
\min \{cx \; : \; x_u + x_v \geq	 1 \text{ for } e=uv \in E, \; x\in [0,1]^{V}\}.
\end{equation}
The integrality gap of this formulation is exactly 2. \cindy{Move to active voice once we have an actual reference for this.} It is shown that it is UG-hard to approximte VC within any factor sctrictly better than 2 \cite{UGhardVC}. \cindy{We compare to small instances} We compare  FDT and the feasbility pump heuristic \cite{fp1} on the PACE 2019\footnote{ Parameterized Algorithms and Computational Experiments Challenge 2019: \url{https://pacechallenge.org/2019/}} challenge test cases \cite{PACE} and compare its performance with 
\subsubsection{Tree augmentation problem}
In the \textsc{Tree Augmentation Problem (TAP)} we are given a  graph $G=(V,E)$, a spanning tree $T$ of $G$. We also have a cost vector $c\in \mathbb{R}^{E\setminus T}_{\geq 0}$. A subset $F$ of $E\setminus T$ is called a \textit{feasible augmentation} if $(V,T\cup F)$ is a 2-edge-connected graph. In TAP we seek the minimum cost feasible augmentation. The natural linear programming relaxation for TAP is 
\begin{equation}
\min \{cx\; : \; \sum_{\ell \in \cov(e)} x_{\ell} \geq 1 \text{ for } e\in T, \; x\in [0,1]^{E\setminus T}\}.
\end{equation}
where $\cov(e)$ is set of edges $\ell \in E\setminus T$ such that $e$ is in the unique cycle of $T\cup \{\ell\}$. We call the LP above the cut-LP. The integrality gap of the cut-LP is known to be between $\frac{3}{2}$ \cite{32gaptap} and $2$ \cite{FJ81}. We create random fractional extreme points of the cut-LP and round them using FDT. For the instances that we create the blow-up factor is always below $\frac{3}{2}$ providing an upper bound for such instances.

@@

Consider a graph $G=(V,E)$. A \textit{Carr-Vempala point} $x\in \mathbb{R}^E$ is a fractional point in $\subtour(G)$ where the edges $e$ with $0<x_e<1$ form a Hamiltonian cycle of $G$. \cindy{I am having a hard time parsing the next sentence.  Is there supposed to be a value for $g(\2ec)$ for this case?} Carr and Vempala \cite{Carr2004} showed that $g(\2ec)$ for instances where the optimal solution to $\min_{x\in \subtour(G)}cx$ is a Carr-Vempala point. We show that the integrality gap is at most $\frac{6}{5}$ for Carr-Vempala points with at most 12 vertices on the Hamiltonian cycle formed by the fractional edges. For Carr-Vempala points we assume that 1-edges are replaced by long paths of 1-edges making these points into potentially harder-to-round instances.
